---
title: "Statistics and R"
author: "Aadam Bodunrin"
date: "13/11/2019"
output: pdf_document
---

## Peter Lanker Practice question

1. Import the Zagat file in R
```{r}
zagat = read.table('Zagat.csv',
                    sep = ',',
                    header = TRUE,
                    stringsAsFactor = TRUE)
```


- a) Create a graph of the empirical distribution function for the variable Price.
```{r}
ec = ecdf(zagat$Price)
plot(ec)
```

- b) What percent of the restaurants in our database have Price value less than or equal to $40?
```{r}
f <- length(subset(zagat$Name, zagat$Price <= 40))
percent <- (f/300)*100
print(percent)
```

- c) Calculate the mean and the standard deviation for the Price variable.
```{r}
m_price = mean(zagat$Price)
print(m_price)
```


```{r}
sd_price = sd(zagat$Price)
print(sd_price)
```


- d) What percent of the Price data falls within 2 standard deviations from the mean?
```{r}
U2=ec(m_price+(2*sd_price))
D2=ec(m_price-(2*sd_price))
SD2 = (U2-D2) * 100
print(SD2)
```

2. In production line 5% of the produced items is defective (typically this proportion is un-
known; we assume it to be known for the sake of this exercise). A quality control inspector
selects a random sample of n = 20 items.

- a) What is the probability that there will be no defective item in the sample?

Probability that an item is defective is 5% (0.05) while probability that an item is not defective is 1 - 0.05 = 0.95 (95%). Therefore the probability that there would be no defective item in the sample is 
```{r}
print(0.95**20)
```

- b) what is the probability that there will be at least 1 defective item in the sample?
The probability that there would be at least 1 defective batch in the sample is
```{r}
print(1-(0.95**20))
```

3. A normal random variable X has mean 3.0 and standard deviation 0.2. What is the
probability that X falls between 2.75 and 3.1?
```{r}
p = pnorm(2.75,mean=3,sd=0.2)
q = pnorm(3.1,mean=3,sd=0.2)
print(q-p)

```


4. Suppose that X follows normal distribution with mean 5.5 and standard deviation 0.3.
Find a number w such that X < w with 30% probability.
```{r}
(qnorm(0.3, mean=5.5, sd=0.3))
```

5. Bluefish purchased at the Lime Beach Fishing Terminal produce a filet weight which
has a mean of 4.5 pounds with a standard deviation of 0.8 pound. If a restaurant manager
purchases 50 such fish, then what is the probability that she will have at least 220 pounds
of filets?

```{r}
q = pnorm(220,mean=4.5,sd=0.8)
print(q)
```

6. A company is interested in estimating μ, the mean number of days of sick leave during
the last year taken by all its employees. They select a random sample of 100 employees and
note the number of sick days taken by each employee in the sample. The following sample
statistics are computed: ¯x = 12.2 days, s = 3 days. Find a 95% confidence interval for μ.
```{r}
m <- 12.2
s <- 3
t = qt(0.975, df = 99)
l= m-t*s/sqrt(100)
u= m+t*s/sqrt(100)
print(l)
print(u)
```

7. A firm that manages rental properties is assessing an expansion into an expensive area
is San Francisco. To cover its costs, the firm needs the average rent in this area to be more
than $1,500 per month. They set up two hypothesis: H0 : μ = 1500 and Ha : μ > 1500.
In order to make a decision the firm obtained rents for a sample of n = 115 rental units in the area. Among these, the average rent is $1,657 with sample standard deviation s = 581.

- a) Test the above hypothesis at 5% significance level.
```{r}
t_stat <- ((1657 - 1500)/581) * sqrt(115)
t_5 <- qt(.975, df = 114)
print(t_stat)
print(t_5)
#SO, we reject H0. The difference between the sample mean $1,657 and $1,500 is statistically different at 5% 
``` 

- b) Calculate the p-value for this test.
```{r}
tval <- pt(t_stat, 114)
print(1 - tval)
```
8. The data file file HEATING deals with the heating bill for dwelling units of various
numbers of rooms. Use R whenever possible in answering the following questions.
```{r}
Heating = read.table('HEATING.csv',
                    sep = ',',
                    header = TRUE,
                    stringsAsFactor = TRUE)
```
- a) Obtain a scatter-plot of the two variables. Which variable should be on the horizontal
axis?
```{r}
plot(Heating$ROOMS, Heating$FUELBILL)
```

- b) Find the linear regression equation resulting from regression of FUELBILL on ROOMS.
Give an interpretation for the slope and the intercept.
```{r}
regHeating = lm(Heating$FUELBILL~Heating$ROOMS)
regHeating

#Slope: For every additional room, you can expect fuelbill to increase by an average of 136.2
#Intercept:If there are no rooms, the expected value for fuelbill is -251.9. The intercept is negative which is hard to since fuelbill is meant to be either 0 or positive
```

- c) Test the hypothesis that the true slope of the regression line is zero.
```{r}
summary(regHeating)
tq <- qt(.975, df = 142)
b1 <- 136.169
sb1 <- 7.098
q <- (b1 + (tq * sb1))
k <- (b1 - (tq * sb1))
print(q)
print(k)
tstat <- b1/sb1
print(tstat)
confint(regHeating, 'Heating$ROOMS', level=0.95)

#We reject the null hypothesis and accept that the slope of the regression line is zero(0).
```


- d) Predict the FUELBILL for a unit with ROOMS=6.
```{r}
estimate <- (-251.9 + (136.2 * 6))
print(estimate)
```

- e) Create a 95% confidence interval for the average FUELBILL of all dwelling units in this
population with ROOMS=6.
```{r}
x = Heating$ROOMS
y = Heating$FUELBILL
new = data.frame(x = 6)
conf = predict(lm(y~x), new, interval="confidence", level = 0.95)
print(conf)

```

- f) Create a 95% prediction interval for a particular dwelling unit with ROOMS variable
equal to 6.
```{r}
pred = predict(lm(y~x), new, interval = "prediction")
pred
```

- g) A particular 6 room unit last year had a heating bill of $958. Do you find this amount
unusually high?

```{r}
x6 <- data.frame(x)
pred=predict(regHeating, x6, interval="prediction")
confid=predict(regHeating, x6, interval="confidence")
plot(x,y)
lines(x,pred[,1], col="blue",lwd=2)
lines(x,pred[,2], col="green",lwd=2)
lines(x,pred[,3], col="green",lwd=2)
lines(x,confid[,2], col="red",lwd=2)
lines(x,confid[,3], col="red",lwd=2)

#Yes!
```
