---
title: "Peter_Lanker"
author: "Aadam"
date: "13/11/2019"
---

## Peter Lanker Practice question

1. Import the Zagat file in R
```{r}
zagat = read.table('Zagat.csv',
                    sep = ',',
                    header = TRUE,
                    stringsAsFactor = TRUE)
```


- a) Create a graph of the empirical distribution function for the variable Price.
```{r}
lm = ecdf(zagat$Price)
plot(lm)
```

- b) What percent of the restaurants in our database have Price value less than or equal to $40?
```{r}
f <- length(subset(zagat$Name, zagat$Price >= 40))
percent <- (f/300)*100
print(percent)
```

- c) Calculate the mean and the standard deviation for the Price variable.
```{r}
m_price = mean(zagat$Price)
print(m_price)
```


```{r}
sd_price = sd(zagat$Price)
print(sd_price)
```


- d) What percent of the Price data falls within 2 standard deviations from the mean?
```{r}
U2=lm(m_price+2*sd_price)
D2=lm(m_price-2*sd_price)
SD2 = (U2-D2) * 100
print(SD2)
```

2. In production line 5% of the produced items is defective (typically this proportion is un-
known; we assume it to be known for the sake of this exercise). A quality control inspector
selects a random sample of n = 20 items.

- a) What is the probability that there will be no defective item in the sample?

Probability that an item is defective is 5% (0.05) while probability that an item is not defective is 1 - 0.05 = 0.95 (95%). Therefore the probability that there would be no defective item in the sample is 
```{r}
print(0.95**20)
```

- b) what is the probability that there will be at least 1 defective item in the sample?
The probability that there would be at least 1 defective batch in the sample is
```{r}
print(1-(0.95**20))
```

3. A normal random variable X has mean 3.0 and standard deviation 0.2. What is the
probability that X falls between 2.75 and 3.1?
```{r}
p = pnorm(2.75,mean=3,sd=0.2)
q = pnorm(3.1,mean=3,sd=0.2)
print(q-p)

```


4. Suppose that X follows normal distribution with mean 5.5 and standard deviation 0.3.
Find a number w such that X < w with 30% probability.
```{r}
(qnorm(0.3, mean=5.5, sd=0.3))
```

5. Bluefish purchased at the Lime Beach Fishing Terminal produce a filet weight which
has a mean of 4.5 pounds with a standard deviation of 0.8 pound. If a restaurant manager
purchases 50 such fish, then what is the probability that she will have at least 220 pounds
of filets?

```{r}
q = pnorm(228,mean=200,sd=40)
print(1-q)
```

6. A company is interested in estimating μ, the mean number of days of sick leave during
the last year taken by all its employees. They select a random sample of 100 employees and
note the number of sick days taken by each employee in the sample. The following sample
statistics are computed: ¯x = 12.2 days, s = 3 days. Find a 95% confidence interval for μ.
```{r}
m <- 12.2
s <- 3
t = qt(0.975, df = 99)
l= m-t*s/sqrt(100)
u= m+t*s/sqrt(100)
print(l)
print(u)
```

7. A firm that manages rental properties is assessing an expansion into an expensive area
is San Francisco. To cover its costs, the firm needs the average rent in this area to be more
than $1,500 per month. They set up two hypothesis: H0 : μ = 1500 and Ha : μ > 1500.
In order to make a decision the firm obtained rents for a sample of n = 115 rental units in the area. Among these, the average rent is $1,657 with sample standard deviation s = 581.

- a) Test the above hypothesis at 5% significance level.
```{r}
t_stat <- ((1657 - 1500)/581) * sqrt(115)
t_5 <- qt(.975, df = 114)
print(t_stat)
print(t_5)
#SO, we reject H0. The difference between the sample mean $1,657 and $1,500 is statistically different at 5% 
``` 

- b) Calculate the p-value for this test.
```{r}
tval <- pt(t_stat, 114)
print(1 - tval)
```
8. The data file file HEATING deals with the heating bill for dwelling units of various
numbers of rooms. Use R whenever possible in answering the following questions.
```{r}
Heating = read.table('HEATING.csv',
                    sep = ',',
                    header = TRUE,
                    stringsAsFactor = TRUE)
```
- a) Obtain a scatter-plot of the two variables. Which variable should be on the horizontal
axis?
```{r}
plot(Heating$ROOMS, Heating$FUELBILL)
```

- b) Find the linear regression equation resulting from regression of FUELBILL on ROOMS.
Give an interpretation for the slope and the intercept.
```{r}
#regHeating = lm(Heating$FUELBILL~Heating$ROOMS)
regHeating = lm(FUELBILL~ROOMS, data = Heating)
regHeating

#Slope: For every additional room, you can expect fuelbill to increase by an average of 136.2
#Intercept:If there are no rooms, the expected value for fuelbill is -251.9. The intercept is negative which is hard to since fuelbill is meant to be either 0 or positive
```

- c) Test the hypothesis that the true slope of the regression line is zero.
```{r}
summary(regHeating)
tq <- qt(.975, df = 142)
b1 <- 136.169
sb1 <- 7.098
q <- (b1 + (tq * sb1))
k <- (b1 - (tq * sb1))
print(q)
print(k)
tstat <- b1/sb1
print(tstat)
confint(regHeating, 'ROOMS', level=0.95)

#We reject the null hypothesis and accept that the slope of the regression line is not zero(0).
```


- d) Predict the FUELBILL for a unit with ROOMS=6.
```{r}
estimate <- (-251.9 + (136.2 * 6))
print(estimate)
```

- e) Create a 95% confidence interval for the average FUELBILL of all dwelling units in this
population with ROOMS=6.
```{r}
x = Heating$ROOMS
y = Heating$FUELBILL
new = data.frame(x = 6)
conf = predict(lm(y~x), new, interval="confidence")
print(conf)

```

- f) Create a 95% prediction interval for a particular dwelling unit with ROOMS variable
equal to 6.
```{r}
pred = predict(lm(y~x), new, interval = "prediction")
pred
```

- g) A particular 6 room unit last year had a heating bill of $958. Do you find this amount
unusually high?

```{r}
plot(x,y)
lines(x,pred[,1], col="blue",lwd=2)
lines(x,pred[,2], col="green",lwd=2)
lines(x,pred[,3], col="green",lwd=2)
lines(x,confid[,2], col="red",lwd=2)
lines(x,confid[,3], col="red",lwd=2)

#Yes!
```

9. You are considering a quality inspection scheme to use on the spark plugs which are sent
from your supplier. These spark plugs come in a shipments of 50,000. Denote the unknown
proportion of defective spark plugs in the shipment by p. Ideally you would like to reject the
shipment if p > .05 and accept it if p ≤ .05. In practice you can’t follow this plan since you
don’t know p. Instead you decide to apply a scheme that consists of the following steps:

A random sample of 20 of the spark plugs will be selected from each shipment. Each of
the selected plugs will be tested to see whether it is defective or not. (The test involves
measuring the plug gap and determining the electrical resistance.) You will note as X the
(random) number of defective plugs in the sample. If X < 2 then the shipment passes your
quality standard. If X ≥ 2 then the shipment fails the quality test and will be returned to
the supplier.

- a) Find the probability that the shipment is rejected when p = .05 (this corresponds to an
“error” since at p = .05 we would want to accept the shipment).


- b) Find the probability that the shipment is accepted when p = .1 (this corresponds to an
“error” again since at p = .1 we would want to reject the shipment).

- c) Find the probability that the shipment is accepted when p = .2.



10. We would like to modify the quality control test described in question 9 above in the
following way. We want to pass the shipment if X < w and reject the shipment when
X ≥ w where w is a number to be determined. Determine the smallest possible value for w
such that the probability of rejecting the shipment when p = .05 is no more than .01, i.e., 1%. 
